---
title: "Sharp nonparametric bounds for decomposition effects with two mediators -- Supplemental Materials"
date: '`r format(Sys.Date(), "%Y-%m-%d")`'
author:
  - Erin E Gabriel
  - Michael C Sachs
  - Arvid Sjölander
output: 
    html_document
---


# Notation and estimands

Here we reproduce the decompositions of the total effect as reported in Daniel et al. 2015, Table 2. Each element of the decomposition is a link to the bounds for that element. The total effect is TE $= p\{Y(1) = 1\} - p\{Y(0) = 1\}$. The terms of the decomposition are defined as follows for $x, x_1, x_2, x_3 \in \{0, 1\}$:

The natural direct effect:
\begin{eqnarray*}
\mbox{NDE}\mbox{-}x_1 x_2 x_3&=&p\{Y(1, M_1(x_1), M_2(x_2, M_1(x_3)))=1\}\\ 
&-&p\{Y(0, M_1(x_1), M_2(x_2, M_1(x_3)))=1\}.
\end{eqnarray*}

The indirect effect through $M_1$ only: 
\begin{eqnarray*}
\mbox{NIE}_{1}\mbox{-}x x_2 x_3 &=&p\{Y(x, M_1(1), M_2(x_2, M_1(x_3)))=1\}\\ &-& p\{Y(x, M_1(0), M_2(x_2, M_1(x_3)))=1\}.
\end{eqnarray*}

The indirect effect through $M_2$ only: 
\begin{eqnarray*}
\mbox{NIE}_{2}\mbox{-}x x_1 x_3 &=&p\{Y(x, M_1(x_1), M_2(1, M_1(x_3)))=1\}\\ &-& p\{Y(x, M_1(x_1), M_2(0, M_1(x_3)))=1\}.
\end{eqnarray*}

The indirect effect through both $M_1$ and $M_2$
\begin{eqnarray*}
\mbox{NIE}_{12}\mbox{-}x x_1 x_2 &=&p\{Y(x, M_1(x_1), M_2(x_2, M_1(1)))=1\}\\ &-& p\{Y(x, M_1(x_1), M_2(x_2, M_1(0)))=1\}.
\end{eqnarray*}


Define the short hand notation for the estimable probabilities as:
$$p_{ym_1m_2\cdot x} = p\{Y=y, M_1=m_1, M_2=m_2|X=x\}.$$ 
For example, $p_{111\cdot 1} = p\{Y=1, M_1=1, M_2=1|X=1\}$.


# Decompositions {#top}

Here we reproduce the decompositions of the total effect as reported in Daniel et al. 2015, Table 2. Each element of the decomposition is a link to the bounds for that element. The terms are defined in the main text. 

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, results = "asis")
```

```{r construct, include = FALSE}
## construct lookup table
## table 2 from Daniel et al 2015
labs <- c("000", "100", "010", "001", "110", "101", "011", "111")
efft <- data.frame(nde = c(1,1,1,1,1,1,2,2,3,3,4,4,5,5,6,6,7,7,8,8,8,8,8,8), 
           nie.1 = c(2,2,5,6,8,8,1,1,5,8,6,8,1,3,1,4,8,8,1,1,3,4,7,7), 
           nie.2 = c(5,8,2,8,2,6,5,8,1,1,8,6,3,1,8,8,1,4,3,7,1,7,1,4), 
           nie.12 = c(8,5,8,2,6,2,8,5,8,6,1,1,8,8,3,1,4,1,7,3,7,1,4,1)
           )
```

```{r table}
for(j in 1:24) {
  
  lntxt <- c("NDE-", "NIE<sub>1</sub>-", "NIE<sub>2</sub>-", "NIE<sub>12</sub>-")
  lktxt <- c("nde.", "nie.1.", "nie.2.", "nie.12.")
  nme <- paste0(lntxt, 
  c(labs[efft$nde[j]], labs[efft$nie.1[j]], 
    labs[efft$nie.2[j]], labs[efft$nie.12[j]]))
  
  lknm <- paste0(lktxt, c(labs[efft$nde[j]], labs[efft$nie.1[j]], 
    labs[efft$nie.2[j]], labs[efft$nie.12[j]]))
  
  ref <- paste0("#", lknm)
  
  cat("TE =", paste(lapply(1:4, function(i) paste0('<a href="', ref[i], '">', nme[i], "</a>")), 
              collapse = " + "))
  cat("</br>")
  
}
```

# Bounds and bounds functions

```{r bnds}
btext_supp <- readRDS("bounds-supp-text.rds")

for(j in 1:8) {
  
 lntxt <- c("NDE-", "NIE<sub>1</sub>-", "NIE<sub>2</sub>-", "NIE<sub>12</sub>-")
  lktxt <- c("nde.", "nie.1.", "nie.2.", "nie.12.")
  
  
  nme <- paste0(lntxt, 
  c(labs[j], labs[j], 
    labs[j], labs[j]))
  
  lknm <- paste0(lktxt, c(labs[j], labs[j], 
    labs[j], labs[j]))
  
  for(i in 1:4) {
    cat(paste0("\n\n## ", nme[i], " {#", lknm[i], "} \n\n"))
    
    cat(gsub("\n", "</br>", gsub("p_\\{([0-1]{3})\\\\cdot ([0-1])\\}", 
             "p<sub>\\1\U00B7\\2</sub>", 
             btext_supp[[lknm[i]]])))
  
    cat(paste0('</br><a href="bndfuncs/', lknm[i], '.rds">Download R function for this bound</a></br>'))
    cat(paste0('<a href="#top">Back to top</a></br></br>'))
  }
}
```


